Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire

### 2016-09-14 manuscript/supplementary.Rnw
Reformat for GCB

\documentclass[9pt,twocolumn,twoside, lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

@@

% Please give the surname of the lead author for the running footer
\leadauthor{Camac} 
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Bayesian models $|$ Biotic interactions $|$ Growth $|$ Recruitment $|$ Mortality}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}
\maketitle

\section*{SI Materials and Methods}
\subsection*{Study site}
Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals\cite{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003\cite{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests\cite{Williams:2008ba}.

@@

In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}. Here, we used fire to create bare ground because Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Westerling:2006ve,Flannigan:2009kl}. Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.

@@

To simulate near-term warmer conditions indicated by the IPCC\cite{IPCC:2013vv}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}. OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm. The chambers were constructed following the International Tundra Experiment (ITEX) protocols\cite{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Fig. S5-7). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsubsection*{Data inputs for growth, mortality and recruitment models}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

@@

The study used 40 open heathland sites\cite{Camac:2013eb} established immediately after the 2003 fires\cite{Williams:2006tl}. All sites have been free of grazing by cattle since 2003. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots~\texttimes~4 transects) per site, subsampling an area of 2000 m$^2$.

Different variables were recorded at the three sampling levels (site, transect and plot). For each site, altitude and TWI were derived from a 30 m resolution digital elevation model\cite{Gallant:2011er} of the Bogong High Plains. TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape\cite{Moore:1993fk}, was derived for each site using R package \texttt{dynatopmodel}\cite{Metcalfe:2016ud}. Across the 40 sites, TWI varied twofold (\Sexpr{min(round(complete_field_survey_data$topo_wetness_index))} to \Sexpr{round(max(complete_field_survey_data$topo_wetness_index))} units; higher units = wetter). Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m, or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}. This altitudinal temperature range is comparable to that observed in the OTC experiment.

The occurrence and severity of fire was determined immediately after the 2003 fires\cite{Williams:2006tl}. The severity of the fires in each burnt site was estimated using mean post-fire twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}. These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals per site. Across the 30 burnt sites, the average site twig diameter ranged from 2 to 13 mm, with smaller diameters assumed to correspond to lower fire intensity. This encompassed the full range of fire severity from lightly scorched sites containing shrubs with only the tips of the outer branches burnt, to severely burnt sites where the majority of large shrub branches were consumed.

\subsubsection*{Data inputs into density and maximum height models}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.

\subsection*{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

@@

where $\widehat{height}_{i,t}$ is the estimated maximum height for individual $i$ at May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error. The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:

@@

We used Bayesian inference and hence needed to specify prior distributions for all model parameters. In both growth models standard deviation parameters (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$, $\sigma_{ind}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25. We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species\cite{Gelman:2006uw}. In both sets of models the intercept terms ($\mu_{spp}^{CTL}$, $\alpha$), coefficients (i.e. inter-tusscok model: $\beta_1$, $\beta_2$ and $\beta_3$) and hyperparameter (i.e. in the multispecies model) for the mean OTC effect, $\mu_{spp}^{OTC}$, were estimated from normal priors centered on zero with a standard deviation of 0.5. We chose this prior as it spanned the range of likely growth trajectories from very slow to very fast growing.

@@

As with growth rates, we modelled each individual's cumulative hazard, \\
$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches and those in inter-tussock gaps. In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the census interval (in years) between censuses, $t$. By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.

@@

Here the standard deviation parameters in both mortality models (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from positive half Cauchy priors with standard deviation of 2.5. The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with standard deviations of 2.5. We also used these Cauchy priors for inter-tussock model coefficients (i.e. $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$. These cauchy priors are the recommended default choice for binomial regression models\cite{Gelman:2008ua}.

@@

\subsubsection*{Modelling drivers of seedling maximum heights}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:

>>>>

\documentclass[9pt,twocolumn,twoside, lineno]{gcb}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{gcbresearcharticle} % Choose template

% Change label of supplementary figures
\renewcommand{\thefigure}{S\arabic{figure}}

@@

% Please give the surname of the lead author for the running footer
\leadauthor{Camac} 
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}
\dates{This manuscript was compiled on \today}
\doi{\url{doi:10.1111/GCB.XXXXX}}

\begin{document}
\suppmaketitle

\section*{SI Materials and Methods}
\subsection*{Study site}
Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals \citep{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.

@@

In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm}. Here, we used fire to create bare ground because Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity \citep{Westerling:2006ve,Flannigan:2009kl}. Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.

@@

To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we used Open Top Chambers (OTCs) \citep{Molau:1996vc}. OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm. The chambers were constructed following the International Tundra Experiment (ITEX) protocols \citep{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig. S5-7). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsubsection*{Data inputs for growth, mortality and recruitment models}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations \citep{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

@@

The study used 40 open heathland sites \citep{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}. All sites have been free of grazing by cattle since 2003. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years \citep{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots~\texttimes~4 transects) per site, subsampling an area of 2000 m$^2$.

Different variables were recorded at the three sampling levels (site, transect and plot). For each site, altitude and TWI were derived from a 30 m resolution digital elevation model \citep{Gallant:2011er} of the Bogong High Plains. TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape \citep{Moore:1993fk}, was derived for each site using R package \texttt{dynatopmodel} \citep{Metcalfe:2016ud}. Across the 40 sites, TWI varied twofold (\Sexpr{min(round(complete_field_survey_data$topo_wetness_index))} to \Sexpr{round(max(complete_field_survey_data$topo_wetness_index))} units; higher units = wetter). Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m, or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation \citep{Brown:1989vx}. This altitudinal temperature range is comparable to that observed in the OTC experiment.

The occurrence and severity of fire was determined immediately after the 2003 fires \citep{Williams:2006tl}. The severity of the fires in each burnt site was estimated using mean post-fire twig diameter, a proxy measure of fire severity \citep{Whight:1999vw}. These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals per site. Across the 30 burnt sites, the average site twig diameter ranged from 2 to 13 mm, with smaller diameters assumed to correspond to lower fire intensity. This encompassed the full range of fire severity from lightly scorched sites containing shrubs with only the tips of the outer branches burnt, to severely burnt sites where the majority of large shrub branches were consumed.

\subsubsection*{Data inputs into density and maximum height models}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations \citep{Gelman:2007te}.

\subsection*{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset \citep{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \citep{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

@@

where $\widehat{height}_{i,t}$ is the estimated maximum height for individual $i$ at May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error. The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation \citep{Tjorve:2010bk}:

@@

We used Bayesian inference and hence needed to specify prior distributions for all model parameters. In both growth models standard deviation parameters (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$, $\sigma_{ind}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25. We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species \citep{Gelman:2006uw}. In both sets of models the intercept terms ($\mu_{spp}^{CTL}$, $\alpha$), coefficients (i.e. inter-tusscok model: $\beta_1$, $\beta_2$ and $\beta_3$) and hyperparameter (i.e. in the multispecies model) for the mean OTC effect, $\mu_{spp}^{OTC}$, were estimated from normal priors centered on zero with a standard deviation of 0.5. We chose this prior as it spanned the range of likely growth trajectories from very slow to very fast growing.

@@

As with growth rates, we modelled each individual's cumulative hazard, \\
$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches and those in inter-tussock gaps. In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the census interval (in years) between censuses, $t$. By doing this our regression model becomes equivalent to a parametric survival analysis \citep{Kleinbaum:2012tn}.

@@

Here the standard deviation parameters in both mortality models (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from positive half Cauchy priors with standard deviation of 2.5. The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with standard deviations of 2.5. We also used these Cauchy priors for inter-tussock model coefficients (i.e. $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$. These cauchy priors are the recommended default choice for binomial regression models \citep{Gelman:2008ua}.

@@

\subsubsection*{Modelling drivers of seedling maximum heights}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C \citep{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:
